### 2021-07-15 src/paper/data/course_of_disease.tex
Hopefully final polishing.

% Susceptibility and Infectiousness
 \comment[id=K]{@Janos: read this.}
The first stage of any disease is the infection. As detailed in
Equation~\ref{eq:infection_probability_formula} the infection
probability depends on the contact type, the calendar date to determine the seasonality,
the age group of the susceptible person and the variant the infectious person is
carrying.
%
The base infection probability of each contact type ($\beta_c$) is estimated inside our
model (Section~\ref{subsec:estimated_params}). How we model and set the seasonality is
detailed in Section~\ref{subsec:seasonality}. For the susceptibility of each age group
($\zeta_a$) we take the estimates of \citet{Davies2020} (Extended Data Fig. 4). Lastly,
we calibrate the infectiousness of the B.1.1.7 variant ($\sigma_{B.1.1.7}$) from
\citet{Davies2021} to 1.67.

@@

 \comment[id=K]{@Janos: Read this!}

Individuals can become immune either through infection ($\gamma_{immune}$) or vaccination
($\gamma_{vacc,\:d}$). As reinfections are estimated to be extremely rare
\citep{AbuRaddad2020}, we simply set the immunity period to one year with probability
one, i.e. everyone that has been infected enjoys immunity for the rest of the simulation
period.

@@

Immunity is binary in our model, i.e. individuals become either sterile immune or remain
susceptible. Thus, we cannot simply use the reported effectiveness but must also include
the risk of asymptomatic and sub-clinical reinfection among the vaccinated in our
probability to become immune upon vaccination. This is important as there is by now ample
evidence that vaccinated individuals can still get infected with SARS-CoV-2 and transmit
the disease \citep{Petter2021, LevineTiefenbrun2021, Harris2021}.

@@

Taking this evidence together we simplify to 35\% of individuals developing immunity 14
days after their vaccination and 40\% after 21 days. For the remaining 25\% we assume
that the vaccine does not protect them. %
This may seem slightly optimistic for how early vaccinations may already have an effect
when looking at infection rates like \citet{Hunter2021}. However, it can be expected that
individuals underestimate the time it takes for the vaccine to work and behave less
carefully after receiving a vaccination. We err on the side of being too optimistic for
the effectiveness of vaccine and assume a significant protection after 14 days as
\citet{LevineTiefenbrun2021} report strongly reduced viral loads already 12 to 28 days
after the first vaccine.

>>>>

% Susceptibility and Infectiousness
The first stage of any disease is the infection. As detailed in
Equation~\ref{eq:infection_probability_formula} the infection
probability depends on the contact type, the calendar date to determine the seasonality,
the age group of the susceptible person and the variant the infectious person is
carrying.
%
The base infection probability of each contact type ($\beta_c$) is estimated inside our
model (Section~\ref{subsec:estimated_params}). How we model and calibrate the seasonality
effect is detailed in Section~\ref{subsec:seasonality}. For the susceptibility of each
age group ($\zeta_a$) we take the estimates of \citet{Davies2020} (Extended Data Fig. 4).
Lastly, we calibrate the infectiousness of the B.1.1.7 variant ($\sigma_{B.1.1.7}$) from
\citet{Davies2021} to 1.67.

@@

Individuals can become immune either through infection ($\gamma_{immune}$) or vaccination
($\gamma_{vacc,\:d}$). As reinfections are very rare \citep{AbuRaddad2020}, we
set the immunity period to one year with probability one, i.e. everyone that has
been infected enjoys immunity for the rest of the simulation period.

@@

Immunity is binary in our model, i.e. individuals achieve either sterile immunity or
remain susceptible. Thus, we cannot simply use the reported effectiveness but must also
include the risk of asymptomatic and sub-clinical reinfection among the vaccinated in our
probability to become immune upon vaccination. This is important as there is ample
evidence by now that vaccinated individuals can still get infected with SARS-CoV-2 and
transmit the disease \citep{Petter2021, LevineTiefenbrun2021, Harris2021}.

@@

To strike a middle ground we assume that 75\% of individuals achieve sterile immunity
after vaccination. This is split into 35\% reaching immunity after 14 days after the
first shot and 40\% reaching immunity after 21 days.

### 2021-07-15 src/paper/data/school_policies.tex
Hopefully final polishing.

% summary
For nurseries, preschools and schools we implement vacations as announced by the German
federal states as well as school closures, emergency care and rotating schooling
schedules where only one half of students attends every other week or day. An
approximation of the share of contacts still taking place with the different school
regulations can be found in Figure~\ref{fig:school_multiplier}. Note that schooling
policies are decided on the state level and usually involved rules that depend on local
incidences. We simplify these rules to one federal policy from the federal incidence and
the policies of the three most populous federal states (North Rhine-Westphalia, Bavaria
and Baden-Württemberg).

% young educ
The policies for preschools and nurseries are similar to the school policies but simpler.
Until November children attend completely normally, starting in November with increased
hygiene measures. Nurseries and preschools stay open until mid December. From mid
December until January 10, nurseries and preschools were nearly completely closed. If
parents could credibly demonstrate that both parents work in systemically relevant
professions and no other child care arrangement was possible, nurseries and preschools
offered so called ``emergency care''. We assume 10\% of children qualified and used
emergency care during this time. After January 10 when more and more parents returned to
work the rules for emergency care were made less strict and we assume a third children
attended to nursery and preschool. This policy stayed in place until February 20.
Afterwards, preschools and nurseries opened normally until mid March. Then during the
third wave the restrictions February were put back into place until end of April when
nurseries and preschools opened again and stayed open \citep{KiTa_BY, KiTa_NRW, KiTa_BW,
KiTa_BWa, KiTa_BYa}.

% school

>>>>

% summary
For nurseries, preschools and schools we implement vacations as announced by the German
federal states as well as school closures, emergency care and rotating schedules where
only one half of students attends every other week or day. An approximation of the share
of contacts still taking place with the different school regulations can be found in
Figure~\ref{fig:school_multiplier}. Note that schooling policies differ between states
and usually involve rules that depend on local incidences. We simplify these rules to one
federal policy from the federal incidence and the policies of the three most populous
federal states (North Rhine-Westphalia, Bavaria and Baden-Württemberg). The testing
policies for schools are described in Section~\ref{subsec:rapid_test_demand}.

% school
Until November schools were open normally. Starting in November, we assume that increased
hygiene measures were taken. Schools stayed open until mid December.
% mid Dec until mid Jan
From mid December until January 10 schools closed and only offered so called ``emergency
care'' for young children whose parents could credibly demonstrate that both had to work
and had no other child care arrangement. Approximately 25\% of primary school children and 5\%
of secondary students attended school as a result.
% Jan and Feb
After January 10 when parents had returned to work the rules for emergency care were
relaxed and approximately a third of primary school children and 10\% of secondary students
attended school as a result. In addition, graduating classes (most adolescents between 16
and 18) were allowed to return to school in a rotating scheme where each class was split
in two groups. Relying on anecdotal evidence we assume that the groups rotate on a daily
basis.% mid Feb to mid March
Starting on February 22 primary school children were also allowed to return to school on
a rotating basis until mid March.
% mid March to Easter
We summarize the school policy from mid March until Easter as all students being on a
rotating school schedule. In addition, children that qualify for emergency care also
attend on days where their group is scheduled to not attend school physically.
% after Easter
After the Easter break schools were mostly closed again. Part of this was a federal law,
the so called ``Bundesnotbremse'' \citep{Notbremse2021} that set rules for schools based
on local incidences that were binding at the time. As a result, most states adjusted
their schooling policies and during April most schools were closed with emergency care
arrangements as in the time from January 10 to February 21. As cases fell schools were
allowed to gradually open. We summarize this as students being on the same rotating
schedule as from mid March to Easter starting on May 1 \citep{schoolBW, schoolBWa,
schoolBWb, schoolBY, schoolBYa, schoolNRW, schoolNRWa, schoolNRWb, schoolNRWc}.

% young educ
The policies for preschools and nurseries are similar to the school policies but simpler.
Until November children attended completely normally, starting in November with increased
hygiene measures. Nurseries and preschools stayed open until mid December. From mid
December until January 10, nurseries and preschools were nearly completely closed. If
parents could credibly demonstrate that both parents work in systemically relevant
professions and no other child care arrangement was possible, nurseries and preschools
offered so called ``emergency care''. We assume 10\% of children qualified and used
emergency care during this time. After January 10 when parents had returned to work the
rules for emergency care were relaxed and we assume a third of children attended nursery
and preschool. This policy stayed in place until February 20. Afterwards, preschools and
nurseries were open normally (maintaining increased hygiene measures) until mid March.
Then during the third wave the restrictions February were put back into place until end
of April when nurseries and preschools opened again and stayed open for the rest of our
simulation period - maintaining increased hygiene measures. \citep{KiTa_BY, KiTa_NRW,
KiTa_BW, KiTa_BWa, KiTa_BYa, schoolBWb}.

### 2021-07-15 src/paper/report.tex
Hopefully final polishing.

At the core of our agent-based model \citep[][we review more literature in Supplementary
    Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts
between heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between an infectious individual and somebody susceptible to
the disease bears the risk of transmitting the virus. Contacts occur in up to four networks:
 \comment[id=J]{I would replace networks by areas because we have more than four networks}
Within the household, at work, at school, or in other settings (leisure activities,
grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
occur at random. Empirical applications can take the population and household structure
from census data and the network-specific frequencies of
contacts from diary data measuring contacts before the pandemic
\citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies depend
on age and geographical location (see Supplementary
Material~\ref{subsec:assortativity}).

@@

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on
age. A possible infection progresses as shown in
Figure~\ref{fig:model_disease_progression}.\comment[id=HM]{New version of figure has lost
the timing. Need to explain d-subscript, maybe even better to add stuff like 1-3 days
back in? Needs to be comprehensible without recurring to appendix} We differentiate
between an initial period of infection without being infectious or showing symptoms,
being infectious (presymptomatic or asymptomatic), showing symptoms, requiring intensive
care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities of
transitioning between these states depend on age; their duration is random and calibrated
to medical literature (for a detailed description see Supplementary
Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
infectiousness is independent of age \citep{Jones2021}.

@@

Two of the most contentious NPIs concern schools and mandates to work from
home. In many countries, schools switched to remote instruction
during the first wave, so did Germany. After the summer break, they were operating at
full capacity with increased hygiene measures, before being closed again from
mid-December onward. Some states started opening them gradually in late February, but
operation at normal capacity did not resume until the beginning of June.
Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
schools starting after Easter, at which point rapid tests had become widely available. We
estimate the realized scenario to have essentially the same effect as a situation with
closed schools. Under fully opened schools with mandatory tests, total infections would
have been 6\% higher; this number rises to 20\% without tests. These effect sizes are
broadly in line with empirical studies \citep[e.g.,][]{Vlachos2021}.\comment[id=HM]{Need
a back-of the envelope calculation} To use another metric, the effective weekly
reproduction number differs by $0.018$ and $0.052$, respectively. In light of the large
negative effects school closures have on children and parents \citep{Luijten2021,
Melegari2021}---and in particular on those with low socio-economic status---these results
in conjunction with hindsight bias suggest that opening schools combined with a testing
strategy would have been beneficial. In other situations, and particular when rapid test
are not available at scale, trade-offs may well be different.

@@

Our analysis has shown that during the transition to high levels of vaccination and
possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail
 \comment[id=J]{Check this number and add to which time period it refers}
prices for rapid tests were below one Euro in early June 2021. While we do not
distinguish between self-administered rapid tests and point of care rapid tests, the
former are likely to play a larger role for indication-driven testing. Widespread
availability at low prices seems important. However, they rely on purely voluntary
participation in a non-public setting. The benefit of point-of-care rapid tests as a
precondition to participate in leisure activities as well as mandatory tests at the
workplace or at school come from screening the entire population. This is important
because disadvantaged groups are less likely to be reached by testing campaigns relying
on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time, these
groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory tests
at school and at the workplace will extend more into these groups. The same goes for
individuals who exhibit a low level of compliance with CoViD-19-related regulations.
Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
it arguably the most effective tool to contain the pandemic in the short run.

>>>>

At the core of our agent-based model \citep[][we review more literature in Supplementary
Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts between
heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
an infectious individual and somebody susceptible to the disease bears the risk of
transmitting the virus. Contacts occur in up to four networks: Within the household, at
work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
depend on age and geographical location (see Supplementary
Material~\ref{subsec:assortativity}).

@@

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities
of transitioning between these states depend on age; their duration is random and
calibrated to medical literature (for a detailed description see Supplementary
Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
infectiousness is independent of age \citep{Jones2021}.

@@

Two of the most contentious NPIs concern schools and mandates to work from home. In many
countries, schools switched to remote instruction during the first wave, so did Germany.
After the summer break, they were operating at full capacity with increased hygiene
measures, before being closed again from mid-December onward. Some states started opening
them gradually in late February, but operation at normal capacity did not resume until
the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of different
policies regarding schools starting after Easter, at which point rapid tests had become
widely available. We estimate the realized scenario to have essentially the same effect
as a situation with closed schools. Under fully opened schools with mandatory tests,
total infections would have been 6\% higher; this number rises to 20\% without tests.
These effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of the
large negative effects school closures have on children and parents \citep{Luijten2021,
Melegari2021}---and in particular on those with low socio-economic status---these results
in conjunction with hindsight bias suggest that opening schools combined with a testing
strategy would have been beneficial. In other situations, and particular when rapid test
are not available at scale, trade-offs may well be different.

@@

Our analysis has shown that during the transition to high levels of vaccination and
possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
estimated to have cost around 50~Euros per capita \citep{Wollmershauser2021}; retail
prices for rapid tests were below one Euro in early June 2021 and below five Euros for
firms. While we do not distinguish between self-administered rapid tests and point of
care rapid tests, the former are likely to play a larger role for indication-driven
testing. Widespread availability at low prices seems important. However, they rely on
purely voluntary participation in a non-public setting. The benefit of point-of-care
rapid tests as a precondition to participate in leisure activities as well as mandatory
tests at the workplace or at school come from screening the entire population. This is
important because disadvantaged groups are less likely to be reached by testing campaigns
relying on voluntary participation \citep[e.g.][]{StillmanTonin2021}; at the same time,
these groups have a higher risk to contract CoViD-19 \citep{KochInstitut2021a}. Mandatory
tests at school and at the workplace will extend more into these groups. The same goes
for individuals who exhibit a low level of compliance with CoViD-19-related regulations.
Compared to vaccinations, rapid testing programmes allow a much quicker roll-out, making
it arguably the most effective tool to contain the pandemic in the short run.
